---
title: "R Notebook"
output: pdf_document
---

```{r}
library(tidyverse)
library(arrow)
library(tidyverse)
library(lubridate)
library(gender)
library(igraph)
library(dplyr)
```


```{r}
applications <- read_parquet("/Users/kaz/DataspellProjects/Org-Analytics/E3/app_data_sample.parquet")
edges <- read_csv("/Users/kaz/DataspellProjects/Org-Analytics/E3/edges_sample.csv")
```

```{r}
library(gender)
examiner_names <- applications %>%
  distinct(examiner_name_first)

examiner_names_gender <- examiner_names %>%
  do(results = gender(.$examiner_name_first, method = "ssa")) %>%
  unnest(cols = c(results), keep_empty = TRUE) %>%
  select(
    examiner_name_first = name,
    gender,
    proportion_female)

# remove extra colums from the gender table
examiner_names_gender <- examiner_names_gender %>%
  select(examiner_name_first, gender)

# joining gender back to the dataset
applications <- applications %>%
  left_join(examiner_names_gender, by = "examiner_name_first")

# cleaning up
rm(examiner_names)
rm(examiner_names_gender)
gc()
```

```{r}

library(wru)

examiner_surnames <- applications %>%
  select(surname = examiner_name_last) %>%
  distinct()

examiner_race <- predict_race(voter.file = examiner_surnames, surname.only = T) %>%
  as_tibble()

examiner_race <- examiner_race %>%
  mutate(max_race_p = pmax(pred.asi, pred.bla, pred.his, pred.oth, pred.whi)) %>%
  mutate(race = case_when(
    max_race_p == pred.asi ~ "Asian",
    max_race_p == pred.bla ~ "black",
    max_race_p == pred.his ~ "Hispanic",
    max_race_p == pred.oth ~ "other",
    max_race_p == pred.whi ~ "white",
    TRUE ~ NA_character_
  ))



# removing extra columns
examiner_race <- examiner_race %>%
  select(surname,race)

applications <- applications %>%
  left_join(examiner_race, by = c("examiner_name_last" = "surname"))

rm(examiner_race)
rm(examiner_surnames)
gc()

```




```{r}
library(lubridate) # to work with dates

examiner_dates <- applications %>%
  select(examiner_id, filing_date, appl_status_date)

examiner_dates <- examiner_dates %>%
  mutate(start_date = ymd(filing_date), end_date = as_date(dmy_hms(appl_status_date)))

examiner_dates <- examiner_dates %>%
  group_by(examiner_id) %>%
  summarise(
    earliest_date = min(start_date, na.rm = TRUE),
    latest_date = max(end_date, na.rm = TRUE),
    tenure_days = interval(earliest_date, latest_date) %/% days(1)
  ) %>%
  filter(year(latest_date)<2018)

applications <- applications %>%
  left_join(examiner_dates, by = "examiner_id")

rm(examiner_dates)
gc()
```



### Pick two art_units
176 and 218
```{r}
workgroup_176 <- applications %>%
        filter(substr(examiner_art_unit, 1, 3) == '176')

workgroup_218 <- applications %>%
        filter(substr(examiner_art_unit, 1, 3) == '218')

# Summary Statistics for Workgroup 367
cat("Summary for Workgroup 176:\n")
workgroup_176 %>% select(c(race, gender)) %>% table() %>% print()

# Summary Statistics for Workgroup 765
cat("\nSummary for Workgroup 218:\n")
workgroup_218 %>% select(c(race, gender)) %>% table() %>% print()
```

### Other Summary Stats
```{r}
summary(workgroup_176 %>% select(tenure_days)) %>% print()
summary(workgroup_218 %>% select(tenure_days)) %>% print()
```

### Plotting - distribution of race
```{r}
# Combine the two groups for plotting, adding a workgroup identifier
applications_combined <- applications %>%
        filter(substr(examiner_art_unit, 1, 3) %in% c('176', '218')) %>%
        mutate(workgroup = case_when(
                substr(examiner_art_unit, 1, 3) == '176' ~ 'Workgroup 176',
                substr(examiner_art_unit, 1, 3) == '218' ~ 'Workgroup 218'
        ))

# Plotting the demographics comparison
ggplot(applications_combined, aes(x = race, fill = workgroup)) +
        geom_bar(position = "dodge") +
        labs(title = "Comparison of Examiner Demographics: Workgroups 176 vs 218",
             x = "Race",
             y = "Count") +
        theme_minimal() +
        scale_fill_manual(values = c("Workgroup 176" = "slateblue", "Workgroup 218" = "darkorange"))
```


### Plotting - distribution of gender
```{r}
applications_combined <- applications %>%
        filter(substr(examiner_art_unit, 1, 3) %in% c('176', '218')) %>%
        mutate(workgroup = case_when(
                substr(examiner_art_unit, 1, 3) == '176' ~ 'Workgroup 176',
                substr(examiner_art_unit, 1, 3) == '218' ~ 'Workgroup 218'
        ))

# Plotting the demographics comparison
ggplot(applications_combined, aes(x = gender, fill = workgroup)) +
        geom_bar(position = "dodge") +
        labs(title = "Comparison of Examiner Demographics: Workgroups 176 vs 218",
             x = "Gender",
             y = "Count") +
        theme_minimal() +
        scale_fill_manual(values = c("Workgroup 176" = "slateblue", "Workgroup 218" = "darkorange"))
```


### Creating advice edge network
```{r}
advice_network <- graph_from_data_frame(d = edges[, c("ego_examiner_id", "alter_examiner_id")], directed = TRUE)
```


### Calculate the centrality measures

```{r}
# Calculate degree centrality for each node (examiner)
degree_centrality <- degree(advice_network, mode = "all")

# Calculate betweenness centrality for each node (examiner)
betweenness_centrality <- betweenness(advice_network, directed = TRUE)

# Create a dataframe of centrality scores
centrality_scores <- data.frame(
  examiner_id = V(advice_network)$name,
  degree = degree_centrality,
  betweenness = betweenness_centrality
)

```

```{r}
workgroup_176$examiner_id <- as.character(workgroup_176$examiner_id)
centrality_scores$examiner_id <- as.character(centrality_scores$examiner_id)

# Merge the centrality scores with the applications data for workgroup 176
applications_176_with_scores <- workgroup_176 %>%
  left_join(centrality_scores, by = "examiner_id")

# Repeat for workgroup 218, ensuring type consistency
workgroup_218$examiner_id <- as.character(workgroup_218$examiner_id)
applications_218_with_scores <- workgroup_218 %>%
  left_join(centrality_scores, by = "examiner_id")
```

### How does between centrality "affect" or correlate with employee characteristic?
```{r}
# Correlation analysis between centrality and tenure_days
cor.test(applications_176_with_scores$degree, applications_176_with_scores$tenure_days, use = "complete.obs")
cor.test(applications_176_with_scores$betweenness, applications_176_with_scores$tenure_days, use = "complete.obs")
```

- slight pos corr

```{r}

# Create a boxplot of betweenness centrality by gender
ggplot(applications_176_with_scores, aes(x = gender, y = degree)) +
  geom_boxplot() +
  labs(title = "Degree Centrality by Gender - 176",
       x = "Gender", y = "Degree Centrality")
```

- Higher degree centrality for female but NA (unidentified gender) is more hihger
> This could indicate several things: if you think about how the gender became unidentified, it could show those groups who
> were not identified are more likely to be more central in the network. This could be a good thing or a bad thing depending on the context.

```{r}
# Create a boxplot of betweenness centrality by gender
ggplot(applications_218_with_scores, aes(x = gender, y = degree)) +
  geom_boxplot() +
  labs(title = "Degree Centrality by Gender - 218",
       x = "Gender", y = "Degree Centrality")
```

- higher degree centrality for female examiners in workgroup 218
- For both, male is low but in this unit, NA is the lowest
-> lower representation of non-US born examiners

**It would be nice to look how these differ by race but it is getting too much for an exercise so i will end here.**