---
title: "Network Analysis and Centrality Measures"
output: md_document
---

```{r setup, include=T}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(igraph)
library(kableExtra)
```

```{r}
# Create a blank matrix
adjacency_matrix <- matrix(0, nrow=10, ncol=10)

# Define the names for the rows and columns for clarity
rownames(adjacency_matrix) <- c("5", "6", "D", "3", "4", "B", "C", "A", "2", "1")
colnames(adjacency_matrix) <- c("5", "6", "D", "3", "4", "B", "C", "A", "2", "1")

# Connections for 5
adjacency_matrix["5", c("D", "6", "3")] <- 1

# Connections for 6
adjacency_matrix["6", c("D", "B", "5")] <- 1

# Connections for D
adjacency_matrix["D", c("5", "6", "3", "B", "C")] <- 1

# Connections for 3
adjacency_matrix["3", c("D", "4", "C", "B", "5")] <- 1

# Connections for 4
adjacency_matrix["4", c("3", "C")] <- 1

# Connections for B
adjacency_matrix["B", c("6", "D", "C", "A", "3")] <- 1

# Connections for C
adjacency_matrix["C", c("3", "4", "B", "D", "A")] <- 1

# Connections for A
adjacency_matrix["A", c("C", "B", "2")] <- 1

# Connections for 2
adjacency_matrix["2", c("A", "1")] <- 1

# Connections for 1
adjacency_matrix["1", "2"] <- 1

# Since the graph is undirected, we mirror the matrix along the diagonal
adjacency_matrix <- adjacency_matrix + t(adjacency_matrix)
# convert where the value is 2 to 1
adjacency_matrix[adjacency_matrix == 2] <- 1

#sort the matrix
adjacency_matrix <- adjacency_matrix[order(rownames(adjacency_matrix)), order(colnames(adjacency_matrix))]

# Print out the matrix
adjacency_matrix
```

```{r}
library(igraph)

graph <- graph_from_adjacency_matrix(adjacency_matrix, mode = "undirected")

# Plot the graph
plot(graph, layout = layout_nicely(graph), vertex.color = "skyblue", edge.arrow.size = 0.5)
```
```{r}
# Calculate degree centrality for all nodes
degree_centrality <- degree(graph, mode = "all")

# Calculate closeness centrality for all nodes
closeness_centrality <- closeness(graph, mode = "all")

# Calculate betweenness centrality for all nodes
betweenness_centrality <- betweenness(graph, directed = FALSE, normalized = TRUE)

centrality_measures_all <- data.frame(
  Degree = degree_centrality,
  Closeness = closeness_centrality,
  Betweenness = betweenness_centrality
)

centrality_measures_all
```


```{r}
# Extract the centralities for seats A-D
seat_choices <- c('A', 'B', 'C', 'D')
centrality_measures <- data.frame(
  Degree = degree_centrality[seat_choices],
  Closeness = closeness_centrality[seat_choices],
  Betweenness = betweenness_centrality[seat_choices]
)
rownames(centrality_measures) <- seat_choices
```


Centrality Measures for Seats A-D

|   | Degree | Closeness | Betweenness |
|---|-------:|----------:|------------:|
| A |      3 |   0.06250 |    0.388889 |
| B |      5 |   0.07143 |    0.250926 |
| C |      5 |   0.07143 |    0.238889 |
| D |      5 |   0.06250 |    0.090741 |




### Seat A
- **Key Feature**: Highest betweenness centrality, lowest degree centrality.
- **Benefit**: Sitting in Seat A positions you as a key bridge in the network, connecting different coworker groups. This is beneficial for meeting diverse individuals --> can offer broad insights into the company culture and cross-departmental projects.
- **Drawback**: The lower degree centrality means you might have fewer direct connections, potentially making it harder to quickly form a close-knit group of friends.

### Seat B
- **Key Feature**: High degree and closeness centrality.
- **Benefit**: access to a larger direct network and easier communication with others due to the high closeness centrality, facilitating exchange of information and support.
- **Drawback**: the focus here is less on being the sole connector and more on being part of a cohesive network, which might **limit exclusive networking**.

### Seat C
- **Key Feature**: Similar to Seat B.

### Seat D
- **Key Feature**: High degree and closeness centrality but lowest betweenness centrality.
- **Benefit**: Similar to B and C
- **Drawback**: The lowest betweenness centrality = less ideal for those looking to play a bridging role between unconnected coworker groups (less networkhub role).

**In Summary**:
- **For broad networking across diverse groups**: **Seat A**
- **For quickly establishing a strong, central presence in your immediate network**: **Seats B, C, and D**


**What if...**
- if you want to make the strongest bond with one person in the bus, where should you sit, assuming every seat is available?
-  1 or 2
- they have the least number of connections, they are far from the rest of the network, and they are not in the middle of the network.

> If you choose 1, then 2 is your only bet and 2 has more "opportunities". So if you want to make the strongest connection, you would sit in seat 2 and only talk to the person in seat 1.


```{r}
degree_centrality <- degree(graph, mode = "all")
closeness_centrality <- closeness(graph, mode = "all")
betweenness_centrality <- betweenness(graph, directed = FALSE)

# Set up the layout of the graph just once to use in all plots
layout <- layout_with_fr(graph)

# Set up the plotting area
par(mfrow = c(2, 2))

# Degree Centrality Plot
plot(graph, layout = layout,
     vertex.label = V(graph)$name,
     vertex.size = degree_centrality * 5,
     vertex.label.cex = 0.8,
     edge.arrow.size = 0.5,
     main = "Degree Centrality")

# Closeness Centrality Plot
plot(graph, layout = layout,
     vertex.label = V(graph)$name,
     vertex.size = closeness_centrality * 500,  # Scale factor to make the sizes visible, adjust as needed
     vertex.label.cex = 0.8,
     edge.arrow.size = 0.5,
     main = "Closeness Centrality")

# Betweenness Centrality Plot
plot(graph, layout = layout,
     vertex.label = V(graph)$name,
     vertex.size = betweenness_centrality / max(betweenness_centrality) * 50,  # Normalize and scale, adjust as needed
     vertex.label.cex = 0.8,
     edge.arrow.size = 0.5,
     main = "Betweenness Centrality")

# Resetting to default single plotting layout
par(mfrow = c(1, 1))
par(mai = c(0, 0, 0.5, 0.5))
```
