---
title: "Exercise1"
output: pdf_document
---

Read the data

```{r}
library(tidyverse)
df <- read_csv("Connections.csv")
```


Get the count of contancts by their current employer + total count
```{r}
count_by_emp <- df %>%
  group_by(Company) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

# total count
total_count <- df %>%
  summarise(count = n())

print(count_by_emp)
print(total_count)
```


Create nodes and edges dataframe to use with igraph (use tidygraph)
```{r}
library(tidygraph)
library(igraph)


# rename columns
df <- df %>%
  rename(First = `First Name`)
df <- df %>%
  rename(Last = `Last Name`)

# create label as a combination of first and first letter of last name
df <- df %>%
  mutate(label = paste(First, substr(Last, 1, 1), sep = " "))

```



If we color "McGill" differently...

```{r}
library(ggraph)

df <- df %>%
        mutate(color = case_when(
                Company == "McGill University" ~ "red",
                Company == "McGill University - Desautels Faculty of Management" ~ "green",
                TRUE ~ "blue"
        ))

nodes <- df %>%
        select(name = label, color)

edges <- expand.grid(from = df$label, to = df$label) %>%
        left_join(df, by = c("from" = "label")) %>%
        rename(from_company = Company) %>%
        left_join(df, by = c("to" = "label")) %>%
        rename(to_company = Company) %>%
        filter(from_company == to_company, from != to) %>%
        select(from, to) %>%
        distinct()


# Use the modified nodes dataframe to create the tbl_graph
graph <- tbl_graph(nodes = nodes, edges = edges, directed = FALSE)

# Plot with color differentiation
ggraph(graph, layout = "fr") +
        geom_edge_link() +
        geom_node_point(aes(color = color), size = 5) +
        geom_node_text(aes(label = name, filter = name == "McGill University"), repel = TRUE) +
        scale_color_identity() +
        theme_void()
```


# What do we see in the network?
It is clear that most of the connections are from McGill University and Desautels Faculty of Management take very concentrated
which means they take a lot of space in my network.


